---
title: "MitoTrace works on single cell SMART-SEQ2 data"
author: "Mingqiang Wang & Lukas Simon"
date: "December 3, 2019"
output: html_document
---

In this analysis, we demonstrate that MitoTrace could identify the germline mutations on single-cell SMART-SEQ2 data whihc publicly available form Darmanis et al. (Darmanis et al. 2017).

Load R libraries
```{r}
library(MitoTrace)
library(pheatmap)
```

Read the bam file, these BAM files can be downloaded from the link: ***.
```{r}
bams <- list.files("/Users/mwang14/Google Drive/00_Texas_Posdoc_Career/01_Zhao/12_Develop_Mitogenotyping_Tools/02_GBM_data/GBM_Bam/", full.names = T, pattern = ".bam$")
```

read human GRCH38 MT genome
```{r}
fasta_loc <- "/Users/mwang14/Google Drive/00_Texas_Posdoc_Career/01_Zhao/12_Develop_Mitogenotyping_Tools/02_GBM_data/GRCH38_MT.fa"
```

calculte the base counts and coverage counts
```{r}
mae_res <- MitoTrace(bam_list = bams, fasta = fasta_loc)
```

Run MitoTrace and calculate allele frequencies
```{r}
af <- calc_allele_frequency(mae_res)
colnames(af) <- unlist(lapply(colnames(af), function(x) { a <- strsplit(x, "\\.") [[1]][1]}))
```

read the patient label
```{r}
ssr_list <- read.table("/Users/mwang14/Google Drive/00_Texas_Posdoc_Career/01_Zhao/12_Develop_Mitogenotyping_Tools/02_GBM_data/rds_all_cell/SRR_Patient_all_list_sorted", sep = "\t")
ok <- intersect(colnames(af), ssr_list$V1)
ssr_list <- ssr_list[match(colnames(af), as.character(ssr_list$V1)),]
```

read the plate relation file
```{r}
plate_relation <- read.table("/Users/mwang14/Google Drive/00_Texas_Posdoc_Career/01_Zhao/12_Develop_Mitogenotyping_Tools/02_GBM_data/rds_all_cell/GBM_SRA_plate_relation", sep = "\t", header = T)
plate_relation <- plate_relation[match(colnames(af), as.character(plate_relation$sra)),]
```

read metadata
```{r}
meta <- read.table("/Users/mwang14/Google Drive/00_Texas_Posdoc_Career/01_Zhao/12_Develop_Mitogenotyping_Tools/02_GBM_data/rds_all_cell/GBM_metadata.csv")
```

read the geo file
```{r}
geo <- read.delim("/Users/mwang14/Google Drive/00_Texas_Posdoc_Career/01_Zhao/12_Develop_Mitogenotyping_Tools/02_GBM_data/rds_all_cell/FromGeoSpreadsheet.tsv", row.names = 1)
geo <- geo[as.character(plate_relation$plate.well),]
```

read the tSNE matrix paper
```{r}
tsne <- read.table("/Users/mwang14/Google Drive/00_Texas_Posdoc_Career/01_Zhao/12_Develop_Mitogenotyping_Tools/02_GBM_data/rds_all_cell/GBM_TSNE.csv")
```

Split by allele frequency
```{r}
asplit <- split(as.character(ssr_list$V1), ssr_list$V2)
sample_af_means <- do.call(cbind, lapply(asplit, function(x){
  subm_af <- af[,x]
  #subm_cov <- cov[,x]
  rowMeans(subm_af, na.rm = T)
}))
cutoff <- 0.75
ok <- which(apply(sample_af_means, 1, function(x) sum(x > cutoff)) > 0)
```

Identify informative germline mutations using AOV
```{r}
pvals <- apply(af[ok,], 1, function(x) try(summary(aov(x ~ geo$V8))[[1]][1, 5]))
ok <- names(which(pvals < 1e-100))
```

generate the heatmap
```{r}
asplit <- split(colnames(af), geo$V8)
samples_ok <- unlist(lapply(asplit, function(x) sample(x, 200)))
indices_ok <- match(samples_ok, colnames(af))
tmp <- af[ok, indices_ok]
anno_col <- data.frame(patient = geo$V8[indices_ok])
rownames(anno_col) <- colnames(tmp)
pheatmap(tmp, show_colnames = F, annotation_col = anno_col, cluster_cols = T)
```

Highlight some sites
```{r}
par(mfrow = c(1,4))
boxplot(split(tmp["11864T>C",], anno_col$patient == "BT_S1"), main = "11864T>C", xlab = "BT_S1", ylab = "Alternative allele frequency")
boxplot(split(tmp["4924G>A",], anno_col$patient == "BT_S2"), main = "4924G>A", xlab = "BT_S2", ylab = "Alternative allele frequency")
boxplot(split(tmp["4790A>G",], anno_col$patient == "BT_S4"), main = "4790A>G", xlab = "BT_S4", ylab = "Alternative allele frequency")
boxplot(split(tmp["709G>A",], anno_col$patient == "BT_S6"), main = "709G>A", xlab = "BT_S6", ylab = "Alternative allele frequency")
```


tSNE visualization of the cell-cell distance matrix by mitochondrial heteroplasmy profiles separates 4 patients.
```{r}
vars <- apply(af, 1, var)
ok <- names(tail(sort(vars), 500))
tmp <- data.matrix(af[ok,])
tmp <- tmp[,which(apply(tmp, 2, var) > 0)]
correl <- cor(tmp)
tData <- Rtsne::Rtsne(1 - abs(correl), perplexity = 20)
plot(tData$Y, col = as.character(meta$Sample.name.color), main = "Patient", pch =20)
```

tSNE visualization of the cell-cell distance matrix by cell type identify
```{r}
tsne <- tsne[rownames(meta),]
plot(tsne, col = as.character(meta$Sample.name.color), pch = 16, xlab = "tSNE1", ylab = "tSNE2")
```

tSNE visualization of the cell cluster by cell type
```{r}
plot(tsne, col = as.character(meta$Cluster_2d_color), pch = 16, xlab = "tSNE1", ylab = "tSNE2")
```





```{r}


```
